Pheromone processing in \emph{Drosophila}

### 2015-04-01 Paper.tex
use T1/utf8
Revert "replace all -> with unicode right arrows"
This reverts commit e9d30d7fcdb1d641fedb2e938e56d6a82ff057aa.
convert -> to rightarrow
* and add textcomp package

%% LyX 2.1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=1.5cm,lmargin=2cm,rmargin=2cm}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{color}
\usepackage{array}
\usepackage{prettyref}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=true,bookmarksopenlevel=1,
 breaklinks=true,pdfborder={0 0 0},backref=false,colorlinks=true]
 {hyperref}
\hypersetup{
 linkcolor=blue, urlcolor=red, citecolor=blue}

@@

Olfactory information carried by PNs converges, apparently randomly,
onto third-order neurons in the MB (Kenyon Cells) \citep{Caron:2013vn}
which integrate inputs from different glomeruli linearly or sublinearly
\citep{Gruntman:2013vn}. This probabilistic wiring is thought to
reflect the MB's role in learning and memory. In contrast, third-order
lateral horn neurons (LHNs) exhibit stereotyped connectivity and odor
responses \citep{Liang:2013nx,Kohl:2013kx,Fisek:2014uq}, in accordance
with the LH's hypothesized role in mediating innate olfactory behaviors.
Since most ORNs are broadly tuned, odorant identity is presumably
encoded by combinatorial activity of ORN ? PN ensembles. While this
distributive model is appropriate for general odor coding (but see
Mansourian \& Stensmyr {[}this issue{]} for a discussion of combinatorial
coding under natural odor concentrations), the only identified olfactory
pheromone, cVA, seems be initially processed by a \emph{labeled line}
instead, i.e. a neural circuit dedicated to a specific sensory stimulus.

@@

How can a single pheromone, cVA, elicit aggression and male-male repulsion
in males \citep{Kurtovic:2007fu,Wang:2010kl,Liu:2011dq}, increase
receptivity in females \citep{Kurtovic:2007fu} and act as aggregation
signal in both sexes \citep{Bartelt:1985bh} (\prettyref{fig:Questions})?
Numerous differences exist between male and female fly brains, most
controlled by \emph{fruitless} \citep{Cachero:2010vn} and \emph{doublesex}
\citep{Rideout:2010dq}. Many of these differences are associated
with sensory processing pathways \citep{Cachero:2010vn}. Males have
a higher number of trichoid sensilla \citep{Stocker:1994fu} and three
enlarged glomeruli, DA1, VA1v and VL2a, all of which express \emph{fru+}
\citep{Stockinger:2005tg}. Deeper in the brain, DA1 PN axons show
a male-specific ventral arborization \citep{Datta:2008dz}, but it
remains unclear if this has functional consequences. Surprisingly,
cVA responses in first- and second-order olfactory neurons are non-dimorphic
\citep{Kurtovic:2007fu,Goes-van-Naters:2007dq,Datta:2008dz}, but
the dendrites of two groups of \emph{fru+} third-order olfactory neurons
sex-specifically overlap with DA1 PN axon terminals, thereby forming
a bidirectional circuit switch that reroutes the cVA signal between
males and females \citep{Kohl:2013kx}. Intriguingly, the position
of this switch is controlled by \emph{fruitless} acting cell-autonomously
in third-order neurons \citep{Kohl:2013kx}. Since only one candidate
fourth-order neuron has been identified so far, the male specific
\emph{fru+} DN1 neuron which projects to the VNC \citep{Ruta:2010ys},
it will now be critical to study how cVA processing after this circuit
bifurcation instructs distinct behaviors. The circuit switch provides
a simple conceptual explanation of how the same sensory stimulus can
elicit sex-specific behaviors, but it does not address (1) how cVA
can act as both repellent and aggregation signal in males and (2)
why prolonged cVA exposure leads to response inhibition in both sexes
\citep{Ejima:2007bs,Liu:2011dq}. As mentioned above, cVA is processed
by two parallel channels \citep{Ha:2006qa,Kurtovic:2007fu,Goes-van-Naters:2007dq}.
While elevated aggression and male-male repulsion are presumably mediated
via acute perception of cVA through the high-affinity Or67d?DA1 channel
\citep{Kurtovic:2007fu,Wang:2010kl}, chronic cVA exposure might recruit
the low-affinity Or65a?DL3 channel and suppress aggression by lateral
inhibition of DA1 in the AL \citep{Ejima:2007bs,Liu:2011dq}. Activation
of Or65a?DL3 (neither of which express \emph{fru}) might therefore
mediate the non-dimorphic aggregation effect of cVA. Lateral inhibition
of DA1 by DL3 might also to be responsible for the abolished attraction
of females to cVA after prolonged exposure \citep{Lebreton:2014cr},
but this interglomerular interaction remains to be directly tested.
Also, in the absence of functional recordings from DL3 PNs or their
postsynaptic neurons, it remains unclear whether additional interactions
exist between both cVA processing channels. Since DL3 PNs project
to the vLH, there are likely to be third-order neurons postsynaptic
to both DA1 and DL3 PNs.

@@

\begin{tabular*}{0.85\paperwidth}{@{\extracolsep{\fill}}|>{\raggedright}p{0.07\paperwidth}|>{\centering}p{0.04\paperwidth}|>{\centering}p{0.06\paperwidth}|>{\centering}p{0.06\paperwidth}|>{\centering}p{0.03\paperwidth}|>{\centering}p{0.12\paperwidth}|>{\centering}p{0.2\paperwidth}|>{\centering}p{0.07\paperwidth}|}
\hline 
\textbf{Pheromone} & \textbf{Range} & \textbf{emitter ? detector} & \textbf{Receptor} & \textbf{PN} & \textbf{Behavior} & \textbf{Comments} & \textbf{References}\tabularnewline
\hline 
\hline 
\multirow{5}{0.07\paperwidth}{cVA} & \multirow{5}{0.04\paperwidth}{volatile \& contact} & M ? M & \textcolor{blue}{Or67d} & \textcolor{blue}{DA1} & aggression & suppressed by chronic cVA via Or65a & \citep{Wang:2010kl,Ejima:2007bs,Liu:2011dq}\tabularnewline
\cline{3-8} 
 &  & F ? M & \textcolor{blue}{Or67d} & \textcolor{blue}{DA1} & male repulsion & transferred to female by prior mating & \citep{Kurtovic:2007fu}\tabularnewline
\cline{3-8} 
 &  & M ? F & \textcolor{blue}{Or67d} & \textcolor{blue}{DA1} & female receptivity & suppressed by chronic cVA via Or65a & \citep{Kurtovic:2007fu,Lebreton:2014cr}\tabularnewline
\cline{3-8} 
 &  & M/F ? M/F & Or65a (?) & DL3 (?) & aggregation\textsuperscript{4} &  & \citep{Bartelt:1985bh}\tabularnewline
\cline{3-8} 
 &  & M ? M & \textcolor{blue}{\emph{ppk23}}\textcolor{blue}{/}\textcolor{blue}{\emph{ppk29}}\textsuperscript{1} & n/a & male-male repulsion & \emph{ppk} channels pheromone receptors ? & \citep{Thistle:2012fu}\tabularnewline
\hline 
\multirow{3}{0.07\paperwidth}{7-T} & \multirow{3}{0.04\paperwidth}{contact} & M ? F & Gr32a & n/a & female receptivity & increases female receptivity & \citep{Grillet:2006tg}\tabularnewline
\cline{3-8} 
 &  & M ? M & Gr32a (Gr66a) & n/a & male-male / interspecies repulsion &  & \citep{Fan:2013fv,Lacaille:2007ij}\tabularnewline
\cline{3-8} 
 &  & M ? M & \textcolor{blue}{\emph{ppk23}}\textcolor{blue}{/}\textcolor{blue}{\emph{ppk29}}\textsuperscript{1} & n/a & male-male repulsion & \emph{ppk} channels pheromone receptors ? & \citep{Toda:2012ff,Starostina:2012mi,Liu:2012ys,Thistle:2012fu}\tabularnewline
\hline 
7-P & contact & M ? F & ? & n/a & female receptivity (?) & male-enriched & \citep{Ferveur:2005ly,Yew:2009fv}\tabularnewline
\hline 
7,11-HD (7,11-ND)\textsuperscript{3} & contact & F ? M & \textcolor{blue}{\emph{ppk23}}\textcolor{blue}{/}\textcolor{blue}{\emph{ppk29}}\textcolor{blue}{/}

\textcolor{blue}{\emph{ppk25}}\textcolor{blue}{/}\textcolor{blue}{\emph{nope}}\textsuperscript{1} & n/a & male courtship & \emph{ppk} channels pheromone receptors ? & \citep{Toda:2012ff,Starostina:2012mi,Liu:2012ys,Thistle:2012fu}\tabularnewline
\hline 
9-P & contact & F ? M & ? & n/a & male copulation & role in courtship conditioning & \citep{Siwicki:2005kl}\tabularnewline
\hline 
\hline 
CH503 & contact (?) & F ? M & ? & ? & male repulsion & transferred to female by prior mating & \citep{Yew:2009fv}\tabularnewline
\hline 
\hline 
9-T, 11-P & contact & M ? M & Gr32a / Gr33a\textsuperscript{5} & n/a & male-male / inter- species repulsion & no cVA-mediated aggression in Gr32a -/- males; contact \emph{fru+}
neurons in GNG & \citep{Svetec:2005kl,Lacaille:2007ij,Miyamoto:2008kl,Moon:2009qa,Koganezawa:2010tg,Fan:2013fv}\tabularnewline
\hline 
\hline 
PAA, PA & volatile & food ? M & \textcolor{blue}{Ir84a} & \textcolor{blue}{VL2a} & male courtship & kairomone aphrodisiac & \citep{Grosjean:2011fu}\tabularnewline
\hline 
\hline 
SP & contact & M ? F & \textcolor{blue}{SPR} & n/a & postmating response & allohormone pheromone\textsuperscript{2}, transferred with seminal
fluid & \citep{Yapici:2008ve}\tabularnewline
\hline 
\hline 
M/F extract & contact & ? & \textcolor{blue}{Or47b} & \textcolor{blue}{VA1v} & ? & active compound(s) unidentified & \citep{Goes-van-Naters:2007dq}\tabularnewline
\hline 
\hline 
M/F extract & contact & ? & \textcolor{blue}{Or88a} & \textcolor{blue}{VA1d} & ? & active compound(s) unidentified & \citep{Goes-van-Naters:2007dq}\tabularnewline
\hline 
? & contact & F ? M & Gr68a & n/a & male courtship & stimulate male courtship, \emph{doublesex}-dependent & \citep{Bray:2003fk}\tabularnewline
\hline 
? & contact & F ? M & Gr39a & n/a & male courtship & role in sustaining male courtship behavior & \citep{Watanabe:2011hc}\tabularnewline
\hline 
? & contact & F ? M & IR52c / IR52d & n/a & male courtship & presumably contact \emph{fru+} neurons in prothoracic ganglia & \citep{Koh:2014cr}\tabularnewline
\hline 
\end{tabular*}
\end{table}

>>>>

%% LyX 2.1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=1.5cm,lmargin=2cm,rmargin=2cm}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{color}
\usepackage{array}
\usepackage{prettyref}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=true,bookmarksopenlevel=1,
 breaklinks=true,pdfborder={0 0 0},backref=false,colorlinks=true]
 {hyperref}
\hypersetup{
 linkcolor=blue, urlcolor=red, citecolor=blue}

@@

Olfactory information carried by PNs converges, apparently randomly,
onto third-order neurons in the MB (Kenyon Cells) \citep{Caron:2013vn}
which integrate inputs from different glomeruli linearly or sublinearly
\citep{Gruntman:2013vn}. This probabilistic wiring is thought to
reflect the MB's role in learning and memory. In contrast, third-order
lateral horn neurons (LHNs) exhibit stereotyped connectivity and odor
responses \citep{Liang:2013nx,Kohl:2013kx,Fisek:2014uq}, in accordance
with the LH's hypothesized role in mediating innate olfactory behaviors.
Since most ORNs are broadly tuned, odorant identity is presumably
encoded by combinatorial activity of ORN → PN ensembles. While this
distributive model is appropriate for general odor coding (but see
Mansourian \& Stensmyr {[}this issue{]} for a discussion of combinatorial
coding under natural odor concentrations), the only identified olfactory
pheromone, cVA, seems be initially processed by a \emph{labeled line}
instead, i.e. a neural circuit dedicated to a specific sensory stimulus.

@@

How can a single pheromone, cVA, elicit aggression and male-male repulsion
in males \citep{Kurtovic:2007fu,Wang:2010kl,Liu:2011dq}, increase
receptivity in females \citep{Kurtovic:2007fu} and act as aggregation
signal in both sexes \citep{Bartelt:1985bh} (\prettyref{fig:Questions})?
Numerous differences exist between male and female fly brains, most
controlled by \emph{fruitless} \citep{Cachero:2010vn} and \emph{doublesex}
\citep{Rideout:2010dq}. Many of these differences are associated
with sensory processing pathways \citep{Cachero:2010vn}. Males have
a higher number of trichoid sensilla \citep{Stocker:1994fu} and three
enlarged glomeruli, DA1, VA1v and VL2a, all of which express \emph{fru+}
\citep{Stockinger:2005tg}. Deeper in the brain, DA1 PN axons show
a male-specific ventral arborization \citep{Datta:2008dz}, but it
remains unclear if this has functional consequences. Surprisingly,
cVA responses in first- and second-order olfactory neurons are non-dimorphic
\citep{Kurtovic:2007fu,Goes-van-Naters:2007dq,Datta:2008dz}, but
the dendrites of two groups of \emph{fru+} third-order olfactory neurons
sex-specifically overlap with DA1 PN axon terminals, thereby forming
a bidirectional circuit switch that reroutes the cVA signal between
males and females \citep{Kohl:2013kx}. Intriguingly, the position
of this switch is controlled by \emph{fruitless} acting cell-autonomously
in third-order neurons \citep{Kohl:2013kx}. Since only one candidate
fourth-order neuron has been identified so far, the male specific
\emph{fru+} DN1 neuron which projects to the VNC \citep{Ruta:2010ys},
it will now be critical to study how cVA processing after this circuit
bifurcation instructs distinct behaviors. The circuit switch provides
a simple conceptual explanation of how the same sensory stimulus can
elicit sex-specific behaviors, but it does not address (1) how cVA
can act as both repellent and aggregation signal in males and (2)
why prolonged cVA exposure leads to response inhibition in both sexes
\citep{Ejima:2007bs,Liu:2011dq}. As mentioned above, cVA is processed
by two parallel channels \citep{Ha:2006qa,Kurtovic:2007fu,Goes-van-Naters:2007dq}.
While elevated aggression and male-male repulsion are presumably mediated
via acute perception of cVA through the high-affinity Or67d→DA1 channel
\citep{Kurtovic:2007fu,Wang:2010kl}, chronic cVA exposure might recruit
the low-affinity Or65a→DL3 channel and suppress aggression by lateral
inhibition of DA1 in the AL \citep{Ejima:2007bs,Liu:2011dq}. Activation
of Or65a→DL3 (neither of which express \emph{fru}) might therefore
mediate the non-dimorphic aggregation effect of cVA. Lateral inhibition
of DA1 by DL3 might also to be responsible for the abolished attraction
of females to cVA after prolonged exposure \citep{Lebreton:2014cr},
but this interglomerular interaction remains to be directly tested.
Also, in the absence of functional recordings from DL3 PNs or their
postsynaptic neurons, it remains unclear whether additional interactions
exist between both cVA processing channels. Since DL3 PNs project
to the vLH, there are likely to be third-order neurons postsynaptic
to both DA1 and DL3 PNs.

@@

\begin{tabular*}{0.85\paperwidth}{@{\extracolsep{\fill}}|>{\raggedright}p{0.07\paperwidth}|>{\centering}p{0.04\paperwidth}|>{\centering}p{0.06\paperwidth}|>{\centering}p{0.06\paperwidth}|>{\centering}p{0.03\paperwidth}|>{\centering}p{0.12\paperwidth}|>{\centering}p{0.2\paperwidth}|>{\centering}p{0.07\paperwidth}|}
\hline 
\textbf{Pheromone} & \textbf{Range} & \textbf{emitter → detector} & \textbf{Receptor} & \textbf{PN} & \textbf{Behavior} & \textbf{Comments} & \textbf{References}\tabularnewline
\hline 
\hline 
\multirow{5}{0.07\paperwidth}{cVA} & \multirow{5}{0.04\paperwidth}{volatile \& contact} & M → M & \textcolor{blue}{Or67d} & \textcolor{blue}{DA1} & aggression & suppressed by chronic cVA via Or65a & \citep{Wang:2010kl,Ejima:2007bs,Liu:2011dq}\tabularnewline
\cline{3-8} 
 &  & F → M & \textcolor{blue}{Or67d} & \textcolor{blue}{DA1} & male repulsion & transferred to female by prior mating & \citep{Kurtovic:2007fu}\tabularnewline
\cline{3-8} 
 &  & M → F & \textcolor{blue}{Or67d} & \textcolor{blue}{DA1} & female receptivity & suppressed by chronic cVA via Or65a & \citep{Kurtovic:2007fu,Lebreton:2014cr}\tabularnewline
\cline{3-8} 
 &  & M/F → M/F & Or65a (?) & DL3 (?) & aggregation\textsuperscript{4} &  & \citep{Bartelt:1985bh}\tabularnewline
\cline{3-8} 
 &  & M → M & \textcolor{blue}{\emph{ppk23}}\textcolor{blue}{/}\textcolor{blue}{\emph{ppk29}}\textsuperscript{1} & n/a & male-male repulsion & \emph{ppk} channels pheromone receptors ? & \citep{Thistle:2012fu}\tabularnewline
\hline 
\multirow{3}{0.07\paperwidth}{7-T} & \multirow{3}{0.04\paperwidth}{contact} & M → F & Gr32a & n/a & female receptivity & increases female receptivity & \citep{Grillet:2006tg}\tabularnewline
\cline{3-8} 
 &  & M → M & Gr32a (Gr66a) & n/a & male-male / interspecies repulsion &  & \citep{Fan:2013fv,Lacaille:2007ij}\tabularnewline
\cline{3-8} 
 &  & M → M & \textcolor{blue}{\emph{ppk23}}\textcolor{blue}{/}\textcolor{blue}{\emph{ppk29}}\textsuperscript{1} & n/a & male-male repulsion & \emph{ppk} channels pheromone receptors ? & \citep{Toda:2012ff,Starostina:2012mi,Liu:2012ys,Thistle:2012fu}\tabularnewline
\hline 
7-P & contact & M → F & ? & n/a & female receptivity (?) & male-enriched & \citep{Ferveur:2005ly,Yew:2009fv}\tabularnewline
\hline 
7,11-HD (7,11-ND)\textsuperscript{3} & contact & F → M & \textcolor{blue}{\emph{ppk23}}\textcolor{blue}{/}\textcolor{blue}{\emph{ppk29}}\textcolor{blue}{/}

\textcolor{blue}{\emph{ppk25}}\textcolor{blue}{/}\textcolor{blue}{\emph{nope}}\textsuperscript{1} & n/a & male courtship & \emph{ppk} channels pheromone receptors ? & \citep{Toda:2012ff,Starostina:2012mi,Liu:2012ys,Thistle:2012fu}\tabularnewline
\hline 
9-P & contact & F → M & ? & n/a & male copulation & role in courtship conditioning & \citep{Siwicki:2005kl}\tabularnewline
\hline 
\hline 
CH503 & contact (?) & F → M & ? & ? & male repulsion & transferred to female by prior mating & \citep{Yew:2009fv}\tabularnewline
\hline 
\hline 
9-T, 11-P & contact & M → M & Gr32a / Gr33a\textsuperscript{5} & n/a & male-male / inter- species repulsion & no cVA-mediated aggression in Gr32a -/- males; contact \emph{fru+}
neurons in GNG & \citep{Svetec:2005kl,Lacaille:2007ij,Miyamoto:2008kl,Moon:2009qa,Koganezawa:2010tg,Fan:2013fv}\tabularnewline
\hline 
\hline 
PAA, PA & volatile & food → M & \textcolor{blue}{Ir84a} & \textcolor{blue}{VL2a} & male courtship & kairomone aphrodisiac & \citep{Grosjean:2011fu}\tabularnewline
\hline 
\hline 
SP & contact & M → F & \textcolor{blue}{SPR} & n/a & postmating response & allohormone pheromone\textsuperscript{2}, transferred with seminal
fluid & \citep{Yapici:2008ve}\tabularnewline
\hline 
\hline 
M/F extract & contact & ? & \textcolor{blue}{Or47b} & \textcolor{blue}{VA1v} & ? & active compound(s) unidentified & \citep{Goes-van-Naters:2007dq}\tabularnewline
\hline 
\hline 
M/F extract & contact & ? & \textcolor{blue}{Or88a} & \textcolor{blue}{VA1d} & ? & active compound(s) unidentified & \citep{Goes-van-Naters:2007dq}\tabularnewline
\hline 
? & contact & F → M & Gr68a & n/a & male courtship & stimulate male courtship, \emph{doublesex}-dependent & \citep{Bray:2003fk}\tabularnewline
\hline 
? & contact & F → M & Gr39a & n/a & male courtship & role in sustaining male courtship behavior & \citep{Watanabe:2011hc}\tabularnewline
\hline 
? & contact & F → M & IR52c / IR52d & n/a & male courtship & presumably contact \emph{fru+} neurons in prothoracic ganglia & \citep{Koh:2014cr}\tabularnewline
\hline 
\end{tabular*}
\end{table}
